---
title: "Up-to-Date-tRicks"
author: "Yaniv"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



***Idea***

I will continue updating this throughout the course. The goal here is a quick cheatsheet for the `R` we have learned or are learning. 

**NOTE 1:** From now, this will grow when 

1.   You provide a usable example OR   
 
2.   You ask a question about a function that you want clarified.   

**NOTE 2:** You should  Also aim to build the skills if googling / finding answers for how to do things in R and how to use `help` files. After all, you don't want to limit yourself to knowing only what I taught.  We will work on those skills together. 

**NOTE 3:** There are numerous excellent R cheat sheets.   

- The RStudio environment https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf 
- Transformation w dplyr https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf 
- Visualization w ggplot2 https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf 
- Using RMarkdown https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf 
- Importing data https://github.com/rstudio/cheatsheets/raw/master/data-import.pdf 


# Intro to R / Fundamentals 

This is associated with work for 1/23 and the introduction to `R` in datacamp  https://www.datacamp.com/courses/free-introduction-to-r  


## Commenting

```{r commenting, results='hold'}
# Commenting. Use the hash aka # to comment. 
# This means R will not try to run these words. 
```

## Arithmetic 

`1 + 1        # Addition ...................` yields: `r 1 + 1`  
`2 * 2        # Multiplication .............` yields: `r 2 * 2`  
`3 ^ 3        # Powers .....................` yields: `r 3 ^ 3`  
`sqrt(4)      # Square roots ...............` yields: `r sqrt(4)`  
`5 - 5        # Subtraction ................` yields: `r 5 - 5`  
`6 / 6        # Division ...................` yields: `r 6 / 6`  
`factorial(7) # Factorial ..................` yields: `r factorial(7)`  
`4 %% 8       # Modulo [example 1] .........` yields: `r 4 %% 8`  
`8 %% 4       # Modulo [example 4] .........` yields: `r 8 %% 4`  



## Assignment 

```{r assignment}
# To keep values in R's head, we need to assign them
# Assignments should show up in RStudio's Environment tab.
# Or you can see all assignments with the ls function
x <- 2       # Assign "x" the value 2
y <- 2
```

`x * 4        # Multiply x by 4 ............` yields: `r x * 4`  
`ls()         # Show whats in Rs head ......` yields: `r ls()`  [in this case]

```{r}
z <- "a"
```

`ls()         # Show whats in Rs head ......` yields: `r ls()`  [in this case]


## Data types 

- Numerics (numbers)
- Logical (TRUE / FALSE)
- Character (words)   
- Factor (words etc that have been stored as integers [this can be convenient or frustrating]). 

Check the classes of different `R` objects:   
`class(1) ..................................` yields: `r class(1)`  
`class(1 > 2) ..............................` yields: `r class(1 > 2)`  
`class("a") ................................` yields: `r class("a")`  
`class(factor("a")) ........................` yields: `r class(factor("a"))`  


## Logical operators

We can ask logical questions, these can be quite useful (for example when we hope to subset our data)  
`x * 4        # Multiply x by 4? ...........` yields: `r x * 4`  
`1   == 2     # Does one equal two? ........` yields: `r 1   == 2`  
`"y" == "x"   # Does "y" equal "x"? ........` yields: `r "y" == "x"`  
`y   == x     # Does y equal x? ............` yields: `r y   == x`  
`1   >  2     # Is one greater than two? ...` yields: `r 1   >  2 `  
`1   <  2     # Is one less than two? ......` yields: `r 1   <  2`  
`1   >= 1     # Is one > than or = to one? .` yields: `r 1   >= 1`  
`1   <= 1     # Is one < than or = to one? .` yields: `r 1   <= 1`  
`1:4 %in% 4:5 # for each value in the vector`  
`1, 2, 3, 4, which are in the vector 4, 5? .` yields: `r 1:4 %in% 4:5`



## Data structures   

Vectors, Matrices, Arrays, Data frames, Tibbles and Lists    
We will focus on vectors and data frames (tibbles are a special kind of data frame). We may come back to lists later. We will almost never use matrices of arrays. 

```{r}
a   <- c(1, 8, 5, 3) # Assign a to a vector w. elements one, eight,  five, and three
b   <- c(john = 1, jen = 8)  
```

`a[3]         # Find the 3rd element in a .` yields: `r a[3]`  
`b["jen"]     # Find jen's value in b .....` yields: `r b["jen"]`  

